---
title: "Project_ExploratoryDataAnalysis"
author: "Marc A.T. Teunis"
date: "5 juni 2016"
output: html_document
---

# Packages used in this assignment
```{r, Packages}
# install.packages("downloader")
# install.packages("dplyr")
# install.packages("lubridate")
library(rmarkdown)
library(downloader)
library(dplyr)
library(lubridate)
```

# Getting the dataset for the project assignment
This assignment uses data from
the <a href="http://archive.ics.uci.edu/ml/">UC Irvine Machine
Learning Repository</a>, a popular repository for machine learning
datasets. In particular, we will be using the "Individual household
electric power consumption Data Set" which I have made available on
the course web site:

## Downloading the zip file, unzipping it and reading the data
```{r, DownloadingData}
url <- "https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2Fhousehold_power_consumption.zip"
download(url, dest="project_dataset.zip", mode="wb") 
unzip("project_dataset.zip", exdir = "./COURSE4EXPLORATORY/data")
```

## Reading the downloaded and unzipped file 
The code above downloads and unzipped the zipped dataset file into the 
data folder ("./COURSE4EXPLORATORY/data").
The code below will point R to that folder to look for the file and read in the data from it. The datafile is called:  "household_powerconsumption.txt"
```{r, ReadingDataSet}
## for help:
# ??read.table
file <- list.files("./COURSE4EXPLORATORY/data", pattern = "txt", full.names = TRUE)
data <- read.table(file, sep = ";", header = TRUE, na.strings = "?")
```

# Selecting the right data 
For the assignment it is necessary to select data only between dates:
2007-02-01 and 2007-02-02
```{r, SelectingCases}
str(data)
tail(data)
## names to lower
names(data) <- tolower(names(data))
names(data)

data$date <- as.character(data$date)
head(data)
data_selected_1 <- data %>% filter(date == "1/2/2007")
data_selected_2 <- data %>% filter(date == "2/2/2007")

data_twoDays <- rbind(data_selected_1,data_selected_2)
head(data_twoDays)
tail(data_twoDays)
```

# Converting the date variable to a data class object
```{r}
data$date <- as.character(data$date)
data$date <- dmy(data$date)
head(data)
str(data)
?lubridate

str(data)
head(data)
```

# Exploring the data 
Generating the graphs for the assignment

```{r, Exploring_Graph1_GlobalActivePower}


 
``` 

