---
title: "Project_ExploratoryDataAnalysis"
author: "Marc A.T. Teunis"
date: "5 juni 2016"
output: html_document
---

# Packages used in this assignment
```{r, Packages}
# install.packages("downloader")
# install.packages("dplyr")
# install.packages("lubridate")
install.packages("ggplot2")

## loading the packages
library(rmarkdown)
library(downloader)
library(dplyr)
library(lubridate)
library(ggplot2)
```

# Getting the dataset for the project assignment
This assignment uses data from
the http://archive.ics.uci.edu/ml/ Irvine Machine
Learning Repository</a>, a popular repository for machine learning
datasets. In particular, we will be using the "Individual household
electric power consumption Data Set" which I have made available on
the course web site:

# Downloading the zip file, unzipping it and reading the data
```{r, DownloadingData}
url <- "https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2Fhousehold_power_consumption.zip"
download(url, dest="project_dataset.zip", mode="wb") 
unzip("project_dataset.zip", exdir = "./data")
```

# Reading the downloaded and unzipped file 
The code above downloads and unzipped the zipped dataset file into the 
data folder ("./data").
The code below will point R to that folder to look for the file and read in the data from it. The datafile is called:  "household_powerconsumption.txt"
```{r, ReadingDataSet}
## for help:
# ??read.table
file <- list.files("./data", pattern = "txt", full.names = TRUE)
data <- read.table(file, sep = ";", header = TRUE, na.strings = "?")
```

# Selecting the right cases 
For the assignment it is necessary to select data only between dates:
2007-02-01 and 2007-02-02
```{r, SelectingCases}
str(data)
tail(data)
## names to lower
names(data) <- tolower(names(data))
names(data)

data$date <- as.character(data$date)
head(data)
data_selected_1 <- data %>% filter(date == "1/2/2007")
data_selected_2 <- data %>% filter(date == "2/2/2007")

data_twoDays <- rbind(data_selected_1,data_selected_2)
head(data_twoDays)
tail(data_twoDays)
```

# Converting the date variable to a date class object "POSIXlt"
```{r}
# converting to "character"
data_twoDays$date <- as.character(data_twoDays$date)
data_twoDays$time <- as.character(data_twoDays$time)

# pasting date time together to be able to create a 
# "POSIXlt" class variable 
data_twoDays$date_time <- paste(data_twoDays$date,
                                data_twoDays$time,
                                sep = " ")
head(data_twoDays)

# converting $date_time to class "POSIXlt"
data_twoDays$date_time <- strptime(data_twoDays$date_time,
                                 "%d/%m/%Y %H:%M:%S")

# inspecting the just created variable "data_twoDays$date_time"
date_time_var <- data_twoDays$date_time
head(date_time_var)
typeof(date_time_var)

POSIXlt

# The POSIXlt data type is a vector, and the entries in the vector have 
# the following meanings:
# [[1]] seconds
# [[2]] minutes
# [[3]] hours
# [[4]] day of month (1-31)
# [[5]] month of the year (0-11)
# [[6]] years since 1900
# [[7]] day of the week (0-6 where 0 represents Sunday)
# [[8]] day of the year (0-365)
# [[9]] Daylight savings indicator (positive if it is daylight 
#       savings)

## Question what day of the week is 1/2/2007?
date_time_var[1][[7]]

## the answer is four
## making a look-up table df
days_of_the_week <- c("Sunday", "Monday", "Thuesday", "Wednesday",
                    "Thursday", "Friday", "Saturday")
POSIXlt_code <- c(0:6)

dow_df <- cbind(days_of_the_week, POSIXlt_code)
dow_df <- as.data.frame(dow_df)
names(dow_df)
which_day <- dow_df %>% filter(POSIXlt_code == 4)

## The final answer:
which_day[1]

```

# Exploring the data
Generating the graph for the assignment "project 1"

## Plot 1
```{r, Graph1_GlobalActivePower}
names(data_twoDays)

## save the file
png(filename = "./images/plot1.png", width = 480, height = 480, units = "px")
hist(data_twoDays$global_active_power, col = "red", xlab = "Global Active Power", ylab = "Frequency", main = "Global Active Power")
dev.off()  
``` 

## Plot 2
```{r, Graph2_GAP_perDay}
levels_date <- levels(as.factor(data_twoDays$date))

## plot2.png in ggplot2 syntax 
png(filename = "./images/plot2.png", width = 480, height = 480, units = "px")
names(data_twoDays)
plot2 <- ggplot(data_twoDays, aes(date_time, global_active_power))
plot2 + geom_line() +
  scale_x_datetime(date_breaks = "1 day", date_labels = c("Saturday", "Thursday", "Friday")) +
  ylab("Global Active Power (kilowatts)") + 
theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 14),
        axis.title.y = element_text(size = 14)) +
  theme(plot.margin = unit(c(1,1,1,1), "cm"))
dev.off()


 +
        margin(t = 2, r = 2, b = 2, l = 2, unit = "pt")
?margin
?scale_x_datetime
??plot.margin

